### 2021-10-13 chapters/introduction.tex
Responses to examiners continued

    The timescale that sets the loss of angular momentum due to gravitational waves is given by~\cite{Paschalidis2012}
    \begin{align}
        t_{\mathrm{GW}} & \approx 150 \left(\frac{M}{2.7\,\msun}\right)^{-1}\left(\frac{R}{10\,\km}\right)^{-2}\left(\frac{f_2}{3.3\,\kHz} \right)^{-4}\left(\frac{e}{0.2} \right)^{-2}\,\mathrm{ms}, \label{eq:Intro:GWAngularMomentumLoss}        
    \end{align}
    where $f_2$ is the dominant post-merger oscillation frequency (see Sec.~\ref{sec:Intro:StrainAndSpectra}) and $e\equiv(a-b)/R$ is the ellipticity of the hypermassive neutron star. 
    Here, $a$ and $b$ are the semi-major and semi-minor axes of the neutron star, respectively, such that $R=(a+b)/2$.
    The reference value, $f_2=3.3\,\kHz$, was inferred from our model in Chapter~\ref{chapter:DetectionPE} for numerical-relativity simulation of SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016}.
    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.\par

    Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
    Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
    Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.
    \redtext{Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.}

@@

    Two models of the time-domain gravitational-wave strain were used for the basis for our model in Chapter~\ref{chapter:DetectionPE}.
    The first model used a third-order exponentially-damped sinusoid and successfully matched a simulation using DD2 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Bauswein2016}.
    The second model implemented a quadratic frequency-drift term over time for $\fpeak$ and used this model  to measure the neutron star radius~\cite{Bose2018}.
    Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
      \par

@@

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

>>>>

    The timescale that sets the loss of angular momentum due to gravitational waves is given by~\cite{Paschalidis2012}
    \begin{align}
        t_{\mathrm{GW}} & \approx 150 \left(\frac{M}{2.7\,\msun}\right)^{-1}\left(\frac{R}{10\,\km}\right)^{-2}\left(\frac{f_2}{3.3\,\kHz} \right)^{-4}\left(\frac{e}{0.2} \right)^{-2}\,\mathrm{ms}, \label{eq:Intro:GWAngularMomentumLoss}        
    \end{align}
    where $f_2$ is the dominant post-merger oscillation frequency (see Sec.~\ref{sec:Intro:StrainAndSpectra}) and $e\equiv(a-b)/R$ is the ellipticity of the hypermassive neutron star. 
    Here, $a$ and $b$ are the semi-major and semi-minor axes of the neutron star, respectively, such that $R=(a+b)/2$.
    The reference value, $f_2=3.3\,\kHz$, was inferred from our model in Chapter~\ref{chapter:DetectionPE} for numerical-relativity simulation of SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016}.
    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.
    \redtext{Numerical-relativity simulations have shown that the gravitational-wave timescale can range from $\mathcal{O}(20)\,\ms$ shortly after merger to $\gtrsim 0.5\,\s$ for longer lived simulations~\cite{Bernuzzi2016,Radice2018}.} \par

    Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
    Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
    Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.
    Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.

@@

    Two models of the time-domain gravitational-wave strain were used for the basis for our model in Chapter~\ref{chapter:DetectionPE}.
    The first model used a third-order exponentially-damped sinusoid and successfully matched a simulation using DD2 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Bauswein2016}.
    The second model implemented a quadratic frequency-drift term over time for $\fpeak$ and used this model  to measure the neutron star radius~\cite{Bose2018}.
    Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
    \redtext{One possible cause of frequency drift would involve the contraction or expansion of the post-merger remnant, which would increase and decrease the frequency, respectively.}
      \par

@@

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests \redtext{(See also Ref.~\cite{Breschi2019} for a discussion on post-merger waveform robustness)}.
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par
